---
title: "STAT380_MP1"
author:
  - name: Isaac Swope
  - name: Andrew
date: "February 22, 2026"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Front Matter
```{r}
# Libraries
library(dplyr)
library(ggplot2)
library(readxl)
library(lubridate)
library(knitr)

# Functions
countNA <- function(col){
  return(sum(is.na(col)))
}

```

## Reading in Data Set
```{r}
games_url <- "https://raw.githubusercontent.com/Swopei/STAT380-MP1_repo/main/data/CODGames2_mp.xlsx"
games_tmp <- tempfile(fileext = ".xlsx")
download.file(games_url, destfile = games_tmp, mode = "wb")
Games <- read_excel(games_tmp)

```

This solution was found using Google Gemini, the prompt used was, "How can a .xlsx file be read into R from an url?"

Gemini generation the following as one possible solution: 

library(readxl)

\# 1. Define the URL
url <- "https://example.com/data.xlsx"

\# 2. Create a temporary file path
dest_file <- tempfile(fileext = ".xlsx")

\# 3. Download the file (CRITICAL: use mode = "wb")
download.file(url, destfile = dest_file, mode = "wb")

\# 4. Read the file
df <- read_excel(dest_file)

\# Optional: clean up the temp file
unlink(dest_file) 


## Data Exploration
This data set contains 27 columns and 211 rows. 

The variable FullPartial is a categorical variable
```{r}
levels(as.factor(Games$FullPartial)) #gets categories
```

FullPartial has two categories: "Full" and "Partial"

```{r}
countNA(as.factor(Games$FullPartial)) #gets NA count
```

FullPartial has no NAs

```{r}
Games_FP <- Games %>%
  group_by(FullPartial) %>%
  summarize(
     count = sum(FullPartial == FullPartial),
     percent = paste0(round(sum(FullPartial == FullPartial)/nrow(Games)*100, 2), "%")
  )

kable(Games_FP)
```


## Proposed Questions


## Question 1: Does Player Performance Change Over Time?


## Question 2: Our Question









